---
title: "IL_2009-2020 Soybeans EDA"
output:
  html_document:
    df_print: paged
---
```{r}
library(xts)
library(forecast)
library(tseries)
library(CausalImpact)
library(TSA)
library(Metrics)
library(tidyverse)
library(dplyr)
```

```{r}
df <- read.csv("C:/Users/SAMSAS/Desktop/Uchi Courses/Barchart/IL_2009-2020.csv")
```

```{r}
df_ts <- df
df_ts$X <- format(as.Date(df_ts$X, "%m/%d/%Y"), "%Y-%m-%d")
df_ts <- xts(df_ts[,-1], order.by = as.Date(df_ts$X))
```

## Compute correlation matrix
```{r}
library(corrplot)

res <- cor(df_ts)
corrplot(res, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)
```
# LST_day
```{r}
LST_day_ts <- df_ts$LST_day
plot(LST_day_ts)
```


# ACF
```{r}
acf(LST_day_ts)
```
# PACF
```{r}
pacf(LST_day_ts)
```
# Dickey-Fuller
```{r}
adf.test(LST_day_ts)
```
Reject the null hypothesis that it is non-stationary

# Decomposition
```{r}
LST_day_ts_yearly <- ts(LST_day_ts, frequency=45.6)
plot(decompose(LST_day_ts_yearly))
```
# MA 10
```{r}
LST_day_ma <- ma(LST_day_ts_yearly, order = 10)
plot(LST_day_ma)
```
# Holt-Winters
```{r}
LST_day_df_yearly <- cbind(df['X'], df['LST_day'])
LST_day_df_yearly.ts <- ts(LST_day_df_yearly$LST_day, frequency = (46)) # Convert df to time series
LST_day_df_yearly.hw <- HoltWinters(x=LST_day_df_yearly.ts) # Run HW
plot(LST_day_df_yearly.hw)
```
# ARIMA
```{r}
LST_day_arima <- auto.arima(LST_day_df_yearly.ts, seasonal = TRUE)

```

```{r}
LST_day_arima
```


# LST_night
```{r}
LST_night_ts <- df_ts$LST_night
plot(LST_night_ts, col = "brown1")
```

```{r}

```


# ACF
```{r}
acf(LST_night_ts)
```
# PACF
```{r}
pacf(LST_night_ts)
```
# Dickey-Fuller
```{r}
adf.test(LST_night_ts)
```
Reject the null hypothesis that it is non-stationary

# Decomposition
```{r}
LST_night_ts_yearly <- ts(LST_night_ts, frequency=45.6)
plot(decompose(LST_night_ts_yearly))
```
# MA 10
```{r}
LST_night_ma <- ma(LST_night_ts_yearly, order = 10)
plot(LST_night_ma)
```
# Holt-Winters
```{r}
LST_night_df_yearly <- cbind(df['X'], df['LST_night'])
LST_night_df_yearly.ts <- ts(LST_night_df_yearly$LST_night, frequency = 45) # Convert df to time series
LST_night_df_yearly.hw <- HoltWinters(x=LST_night_df_yearly.ts) # Run HW
plot(LST_night_df_yearly.hw)
```
# ARIMA
```{r}
LST_night_arima <- auto.arima(LST_night_df_yearly.ts, seasonal = TRUE)
LST_night_arima
```
```{r}
plot(forecast(LST_night_arima))
```


## NDWI
```{r}
NDWI_ts <- df_ts$NDWI
plot(NDWI_ts)
```
# ACF
```{r}
acf(NDWI_ts)
```
# PACF
```{r}
pacf(NDWI_ts)
```
# Dickey-Fuller
```{r}
adf.test(NDWI_ts)
```
Reject the null hypothesis that it is non-stationary

# Decomposition
```{r}
NDWI_ts_yearly <- ts(NDWI_ts, frequency=45.6)
plot(decompose(NDWI_ts_yearly))
```
# MA 10
```{r}
NDWI_ma <- ma(NDWI_ts_yearly, order = 10)
plot(NDWI_ma)
```
# Holt-Winters
```{r}
NDWI_df_yearly <- cbind(df['X'], df['NDWI'])
NDWI_df_yearly.ts <- ts(NDWI_df_yearly$NDWI, frequency = 46) # Convert df to time series
NDWI_df_yearly.hw <- HoltWinters(NDWI_df_yearly.ts) # Run HW
plot(NDWI_df_yearly.hw)
```
# ARIMA
```{r}
NDWI_arima <- auto.arima(NDWI_df_yearly.ts, seasonal = TRUE)
NDWI_arima
```


```{r}
plot(forecast(NDWI_arima))
```

## NDVI
```{r}
NDVI_ts <- df_ts$NDVI
plot(NDVI_ts)
```
# ACF
```{r}
acf(NDVI_ts)
```
# PACF
```{r}
pacf(NDVI_ts)
```
# Dickey-Fuller
```{r}
adf.test(NDVI_ts)
```
Reject the null hypothesis that it is non-stationary

# Decomposition
```{r}
NDVI_ts_yearly <- ts(NDVI_ts, frequency=45.6)
plot(decompose(NDVI_ts_yearly))
```
# MA 10
```{r}
NDVI_ma <- ma(NDVI_ts_yearly, order = 10)
plot(NDVI_ma)
```
# Holt-Winters
```{r}
NDVI_df_yearly <- cbind(df['X'], df['NDVI'])
NDVI_df_yearly.ts <- ts(NDVI_df_yearly$NDVI, frequency = 46) # Convert df to time series
NDVI_df_yearly.hw <- HoltWinters(x=NDVI_df_yearly.ts) # Run HW
plot(NDVI_df_yearly.hw)
```
# ARIMA
```{r}
NDVI_arima <- auto.arima(NDVI_df_yearly.ts, seasonal = TRUE)
NDVI_arima
```

```{r}
plot(forecast(NDVI_arima),h=46)
```


## TMAX
```{r}
TMAX_ts <- df_ts$TMAX
plot(TMAX_ts)
```
# ACF
```{r}
acf(TMAX_ts)
```
# PACF
```{r}
pacf(TMAX_ts)
```
# Dickey-Fuller
```{r}
adf.test(TMAX_ts)
```
Reject the null hypothesis that it is non-stationary

# Decomposition
```{r}
TMAX_ts_yearly <- ts(TMAX_ts, frequency=45.6)
plot(decompose(TMAX_ts_yearly))
```
# MA 10
```{r}
TMAX_ma <- ma(TMAX_ts_yearly, order = 10)
plot(TMAX_ma)
```
# Holt-Winters
```{r}
TMAX_df_yearly <- cbind(df['X'], df['TMAX'])
TMAX_df_yearly.ts <- ts(TMAX_df_yearly$TMAX, frequency = 45) # Convert df to time series
TMAX_df_yearly.hw <- HoltWinters(x=TMAX_df_yearly.ts) # Run HW
plot(TMAX_df_yearly.hw)
```
# ARIMA
```{r}
TMAX_arima <- auto.arima(TMAX_df_yearly.ts, seasonal = TRUE)
TMAX_arima
```
```{r}
plot(forecast(TMAX_arima))
```

## TMIN
```{r}
TMIN_ts <- df_ts$TMIN
plot(TMIN_ts)
```
# ACF
```{r}
acf(TMIN_ts)
```
# PACF
```{r}
pacf(TMIN_ts)
```
# Dickey-Fuller
```{r}
adf.test(TMIN_ts)
```
Reject the null hypothesis that it is non-stationary

# Decomposition
```{r}
TMIN_ts_yearly <- ts(TMIN_ts, frequency=45.6)
plot(decompose(TMIN_ts_yearly))
```
# MA 10
```{r}
TMIN_ma <- ma(TMIN_ts_yearly, order = 10)
plot(TMIN_ma)
```
# Holt-Winters
```{r}
TMIN_df_yearly <- cbind(df['X'], df['TMIN'])
TMIN_df_yearly.ts <- ts(TMIN_df_yearly$TMIN, frequency = 45) # Convert df to time series
TMIN_df_yearly.hw <- HoltWinters(x=TMIN_df_yearly.ts) # Run HW
plot(TMIN_df_yearly.hw)
```
# ARIMA
```{r}
TMIN_arima <- auto.arima(TMIN_df_yearly.ts, seasonal=TRUE)
TMIN_arima
```
```{r}
plot(forecast(TMIN_arima))
```

```{r}
Omega(LST_day_df_yearly.ts, spectrum.control = list(method = "pgram"))
```

```{r}
Omega(LST_night_df_yearly.ts, spectrum.control = list(method = "pgram"))
```

```{r}
Omega(TMAX_df_yearly.ts, spectrum.control = list(method = "pgram"))
```

```{r}
Omega(TMIN_df_yearly.ts, spectrum.control = list(method = "pgram"))
```

```{r}
Omega(NDVI_df_yearly.ts, spectrum.control = list(method = "pgram"))
```

```{r}
Omega(NDWI_df_yearly.ts, spectrum.control = list(method = "pgram"))
```

```{r}
Omega(PRCP_df_yearly.ts, spectrum.control = list(method = "pgram"))
```

## PRCP
```{r}
PRCP_ts <- df_ts$PRCP
plot(PRCP_ts)
```
# ACF
```{r}
acf(PRCP_ts)
```
# PACF
```{r}
pacf(PRCP_ts)
```
# Dickey-Fuller
```{r}
adf.test(PRCP_ts)
```
Reject the null hypothesis that it is non-stationary

# Decomposition
```{r}
PRCP_ts_yearly <- ts(PRCP_ts, frequency=45.6)
plot(decompose(PRCP_ts_yearly))
```
# MA 10
```{r}
PRCP_ma <- ma(PRCP_ts_yearly, order = 10)
plot(PRCP_ma)
```
# Holt-Winters
```{r}
PRCP_df_yearly <- cbind(df['X'], df['PRCP'])
PRCP_df_yearly.ts <- ts(PRCP_df_yearly$PRCP, frequency = 45) # Convert df to time series
PRCP_df_yearly.hw <- HoltWinters(x=PRCP_df_yearly.ts) # Run HW
plot(PRCP_df_yearly.hw)
```
# ARIMA
```{r}
PRCP_arima <- auto.arima(PRCP_df_yearly.ts, seasonal = TRUE)
PRCP_arima
```
```{r}
plot(forecast(PRCP_arima))
```
```{r}
plot(forecast(PRCP_df_yearly.hw))
```

```{r}
plot(LST_day_ts, col = "brown1")
lines(LST_night_ts, col = "deepskyblue")
```

```{r}
plot(TMAX_ts, col = "brown1")
lines(TMIN_ts, col = "deepskyblue")
```

```{r}
plot(forecast(LST_day_df_yearly.hw))
```
```{r}
plot(forecast(LST_night_df_yearly.hw))
```
```{r}
plot(forecast(TMAX_df_yearly.hw))
```
```{r}
plot(forecast(TMIN_df_yearly.hw))
```
```{r}
plot(forecast(NDWI_df_yearly.hw))
```
```{r}
plot(forecast(NDVI_df_yearly.hw))
```
```{r}
LST_day_arima1 <- auto.arima(LST_day_ts)
plot(forecast(LST_day_arima1,h = 46))
```
```{r}
periodogram(LST_day_ts,ylab='LST_day_ts Periodogram')
```
```{r}
periodogram(LST_night_ts,ylab='LST_night_ts Periodogram')
```
```{r}
periodogram(TMAX_ts,ylab='TMAX_ts Periodogram')
```
```{r}
periodogram(TMIN_ts,ylab='TMIN_ts Periodogram')
```

```{r}
p_ndwi <- periodogram(NDWI_ts,ylab='NDWI_ts Periodogram')
```
```{r}
dd = data.frame(freq=p_ndwi$freq, spec=p_ndwi$spec)
order = dd[order(-dd$spec),]
top2 = head(order, 2)
 
# display the 2 highest "power" frequencies
top2$period <- 1/top2$freq 

top2
```

```{r}
ndwi_temp <- msts(NDWI_ts, seasonal.periods=c(44, 23))
model_ndwi <- tbats(ndwi_temp)
comp <- tbats.components(model_ndwi)
plot(comp)
```
```{r}
plot(forecast(model_ndwi))
```


```{r}
p_ndvi <- periodogram(NDVI_ts,ylab='NDVI_ts Periodogram')
```
```{r}
dd = data.frame(freq=p_ndvi$freq, spec=p_ndvi$spec)
order = dd[order(-dd$spec),]
top2 = head(order, 2)
 
# display the 2 highest "power" frequencies
top2$period <- 1/top2$freq 

top2
```

```{r}
ndvi_temp <- msts(NDVI_ts, seasonal.periods=c(44, 48))
model_ndvi <- tbats(ndvi_temp)
comp <- tbats.components(model_ndvi)
plot(comp)
```
```{r}
plot(forecast(model_ndvi))
```

```{r}
p_PRCP <- periodogram(PRCP_ts,ylab='NDWI_ts Periodogram')
```

```{r}
dynamic_reg_Fourier_Arima <- auto.arima(NDWI_df_yearly.ts,xreg = fourier(ts(NDWI_df_yearly.ts,frequency = 45),K=2))
summary(dynamic_reg_Fourier_Arima)
```

```{r}
NDWI_fcast <- forecast(dynamic_reg_Fourier_Arima, xreg=fourier(NDWI_df_yearly.ts, K=2, 
h=46))
plot(NDWI_fcast)
```

```{r}
dynamic_reg_Fourier_Arima2 <- auto.arima(NDVI_df_yearly.ts,xreg = fourier(ts(NDVI_df_yearly.ts,frequency = 46),K=2))
summary(dynamic_reg_Fourier_Arima2)
```

```{r}
NDVI_fcast <- forecast(dynamic_reg_Fourier_Arima2, xreg=fourier(NDVI_df_yearly.ts, K=2, 
h=46))
plot(NDVI_fcast)
```

## sMAPE comparison

```{r}
df_yield <- read.csv('C:/Users/SAMSAS/Desktop/Uchi Courses/Time Series/IL_soybean_2009_2021.csv')

df_features <- read.csv('C:/Users/SAMSAS/Desktop/Uchi Courses/Time Series/IL_2009-2021.csv')
```

```{r}
df_features_2021 <- xts(df_features[,-1], order.by = as.Date(df_features$X))

df_features_2021 <- df_features_2021["2021-01-02
/2021-12-28"]
```

```{r}
LST_day_sarima_forecast <- forecast(LST_day_arima, h = 46)
LST_night_sarima_forecast <- forecast(LST_night_arima, h = 46)
TMAX_sarima_forecast <- forecast(TMAX_arima, h = 46)
TMIN_sarima_forecast <- forecast(TMIN_arima, h = 46)
NDWI_sarima_forecast <- forecast(NDWI_arima, h = 46)
NDVI_sarima_forecast <- forecast(NDVI_arima, h = 46)
PRCP_sarima_forecast <- forecast(PRCP_arima, h = 46)
```

```{r}

LST_day_hw_forecast <- forecast(LST_day_df_yearly.hw, h = 46)
LST_night_hw_forecast <- forecast(LST_night_df_yearly.hw, h = 46)
TMAX_hw_forecast <- forecast(TMAX_df_yearly.hw, h = 46)
TMIN_hw_forecast <- forecast(TMIN_df_yearly.hw, h = 46)
NDWI_hw_forecast <- forecast(NDWI_df_yearly.hw, h = 46)
NDVI_hw_forecast <- forecast(NDVI_df_yearly.hw, h = 46)
PRCP_hw_forecast <- forecast(PRCP_df_yearly.hw, h = 46)
```

```{r}
LST_day_ms_forecast <- "NA"
LST_night_ms_forecast <- "NA"
TMAX_ms_forecast <- "NA"
TMIN_ms_forecast <- "NA"
NDWI_ms_forecast <- forecast(model_ndwi, h = 46)
NDVI_ms_forecast <- forecast(model_ndvi, h = 46)
PRCP_ms_forecast <- "NA"
```

```{r}
LST_day_fourier_forecast <- "NA"
LST_night_fourier_forecast <- "NA"
TMAX_fourier_forecast <- "NA"
TMIN_fourier_forecast <- "NA"
NDWI_fourier_forecast <- forecast(NDWI_fcast, h = 46)
NDVI_fourier_forecast <- forecast(NDVI_fcast, h = 46)
PRCP_fourier_forecast <- "NA"
```


```{r}
LST_day_sarima_smape <- rmse(df_features_2021$LST_day, LST_day_sarima_forecast$mean)
LST_night_sarima_smape <- rmse(df_features_2021$LST_night, LST_night_sarima_forecast$mean)
TMAX_sarima_smape <- rmse(df_features_2021$TMAX, TMAX_sarima_forecast$mean)
TMIN_sarima_smape <- rmse(df_features_2021$TMIN, TMIN_sarima_forecast$mean)
NDWI_sarima_smape <- rmse(df_features_2021$NDWI, NDWI_sarima_forecast$mean)
NDVI_sarima_smape <- rmse(df_features_2021$NDVI, NDVI_sarima_forecast$mean)
PRCP_sarima_smape <- rmse(df_features_2021$PRCP, PRCP_sarima_forecast$mean)
```

```{r}
LST_day_hw_smape <- rmse(df_features_2021$LST_day, LST_day_hw_forecast$mean)
LST_night_hw_smape <- rmse(df_features_2021$LST_night, LST_night_hw_forecast$mean)
TMAX_hw_smape <- rmse(df_features_2021$TMAX, TMAX_hw_forecast$mean)
TMIN_hw_smape <- rmse(df_features_2021$TMIN, TMIN_hw_forecast$mean)
NDWI_hw_smape <- rmse(df_features_2021$NDWI, NDWI_hw_forecast$mean)
NDVI_hw_smape <- rmse(df_features_2021$NDVI, NDVI_hw_forecast$mean)
PRCP_hw_smape <- rmse(df_features_2021$PRCP, PRCP_hw_forecast$mean)
```

```{r}
LST_day_ms_smape <- "NA"
LST_night_ms_smape <- "NA"
TMAX_ms_smape <- "NA"
TMIN_ms_smape <- "NA"
NDWI_ms_smape <- rmse(df_features_2021$NDWI, NDWI_ms_forecast$mean)
NDVI_ms_smape <- rmse(df_features_2021$NDVI, NDVI_ms_forecast$mean)
PRCP_ms_smape <- "NA"
```

```{r}
LST_day_fourier_smape <- "NA"
LST_night_fourier_smape <- "NA"
TMAX_fourier_smape <- "NA"
TMIN_fourier_smape <- "NA"
NDWI_fourier_smape <- rmse(df_features_2021$NDVI, NDWI_fcast$mean)
NDVI_fourier_smape <- rmse(df_features_2021$NDVI, NDVI_fcast$mean)
PRCP_fourier_smape <- "NA"
```

```{r}
LST_day <- c(LST_day_sarima_smape, LST_day_hw_smape, LST_day_ms_smape, LST_day_fourier_smape)
LST_night <- c(LST_night_sarima_smape, LST_night_hw_smape, LST_night_ms_smape, LST_night_fourier_smape)
TMAX <- c(TMAX_sarima_smape, TMAX_hw_smape, TMAX_ms_smape, TMAX_fourier_smape)
TMIN <- c(TMIN_sarima_smape, TMIN_hw_smape, TMIN_ms_smape, TMIN_fourier_smape)
NDWI <- c(NDWI_sarima_smape, NDWI_hw_smape, NDWI_ms_smape, NDWI_fourier_smape)
NDVI <- c(NDVI_sarima_smape, NDVI_hw_smape, NDVI_ms_smape, NDVI_fourier_smape)
PRCP <- c(PRCP_sarima_smape, PRCP_hw_smape, PRCP_ms_smape, PRCP_fourier_smape)
```

```{r}
df_result <- data.frame(LST_day, LST_night, TMAX, TMIN, NDWI, NDVI, PRCP)
rownames(df_result) <- c("SARIMA","Holt-Winter", "Multiple Seasonal", "Fourier Arima")
```

```{r}
df_result
```
```{r}
PRCP_2021 <- df_features_2021$PRCP
PRCP_2021 <- ts(PRCP_2021, frequency = 1)
plot(PRCP_2021, col="black")
lines(PRCP_fcast, col="red")
lines(ts(PRCP_sarima_forecast$mean, frequency = 1), col="blue")

```
```{r}
PRCP_fcast <- PRCP_hw_forecast$mean
PRCP_fcast <- as.data.frame(PRCP_fcast)
plot(PRCP_fcast)
```
```{r}
grep("actual_yield", colnames(df_2021))
```
```{r}
growing_season <- df_2021[,122:199]
growing_season <- scale(growing_season,center=TRUE,scale=TRUE)
growing_season <- cbind(growing_season, df_2021$actual_yield)
growing_season <- as.data.frame(growing_season)
```

```{r}
result <- c()
for (i in 1:77) {
  fit <- lm(df_2021$actual_yield~growing_season[,i], data = growing_season)
  result <- rbind(result, c(colnames(growing_season)[i], coef(summary(fit))[2,c(1,2)]))
}
```

```{r}
result <- as.data.frame(result)
estimate <- as.numeric(result$Estimate)
estimate
```
```{r}
PRCP_sarima_forecast
```

```{r}
LST_day_growing_season <- LST_day_sarima_forecast$mean[17:27]
LST_night_growing_season <- LST_day_sarima_forecast$mean[17:27]
NDWI_growing_season <- NDWI_sarima_forecast$mean[17:27]
NDVI_growing_season <- NDVI_sarima_forecast$mean[17:27]
TMAX_growing_season <- TMAX_sarima_forecast$mean[17:27]
TMIN_growing_season <- TMIN_sarima_forecast$mean[17:27]
PRCP_growing_season <- PRCP_sarima_forecast$mean[17:27]

LST_day_growing_season <- scale(LST_day_growing_season,center=TRUE,scale=TRUE)
LST_night_growing_season <- scale(LST_night_growing_season,center=TRUE,scale=TRUE)
NDWI_growing_season <- scale(NDWI_growing_season,center=TRUE,scale=TRUE)
NDVI_growing_season <- scale(NDVI_growing_season,center=TRUE,scale=TRUE)
TMAX_growing_season <- scale(TMAX_growing_season,center=TRUE,scale=TRUE)
TMIN_growing_season <- scale(TMIN_growing_season,center=TRUE,scale=TRUE)
PRCP_growing_season <- scale(PRCP_growing_season,center=TRUE,scale=TRUE)

repeat1 <- c()

for (i in 1:11) {
  repeat1 <- append(repeat1, LST_day_growing_season[i])
  repeat1 <- append(repeat1, LST_night_growing_season[i])
  repeat1 <- append(repeat1, NDWI_growing_season[i])
  repeat1 <- append(repeat1, NDVI_growing_season[i])
  repeat1 <- append(repeat1, TMAX_growing_season[i])
  repeat1 <- append(repeat1, TMIN_growing_season[i])
  repeat1 <- append(repeat1, PRCP_growing_season[i])
}
```

```{r}
repeat1
```
```{r}
LST_day_sarima_forecast
```

```{r}
predicted_result <- as.list(repeat1 * estimate)
```

```{r}
yield <- 0
for (i in predicted_result) {
  yield = yield + i
}

yield
```
```{r}
state_yield_2021 = df_2021 %>% group_by(year) %>% summarise_each(funs(Mean = mean))
state_yield_2021 <- state_yield_2021[,ncol(state_yield_2021)-2]
```

```{r}
state_yield_2021
```



